The reproductive and parental care transcriptome of the rock dove

### 2020-09-08 manuscript/elife-v1.tex
now with 3 awesome tables

We created the rock dove transcriptome by sequenced 987 RNA samples from
the hypothalamus, pituitary and gonads from the female and male of 166
rock dove pairs across (Fig. 1). These samples come from nine timepoints
that encapsulate important transitions in reproduction (Fig. 1A), four
offspring removal manipulations (Fig. 1B), and three offspring
replacement manipulations (Fig. 1C) to determine the influence of
internal mechanisms and external stimuli on gene expression. Using
T-distributed Stochastic Neighbor Embedding (tSNE) we show broad
patterns of variation due to treatment or sex in each tissue (Fig. 1D).
In the hypothalamus, tSNE dimension visually separates most of the
samples that were subject to an offspring manipulation from those from
naturally occurring parental stages. In the pituitary however, samples
cluster more by wether they were before incubation day 9 or after
incubation day 17. Neither the hypothalamus or pituitary show strong sex
differences at this level, but the gonads samples from very distinct,
sex-specific clusters.

@@

\caption{\label{tab:table1}Changes in candidate gene expression. Significant chagns in gene expression are shown for X candidate genes. Positive and negative changes are denoted with (+) and (-) respectively for female (F) and male (M) hypothalamus (H), pituitary (P) and gonads (G). Comparisons are denoted in the column names. The first group is the reference group.}
\centering
\begin{tabular}[t]{l|l|l|l|l|l|l|l}
\hline
gene & bldg-lay & lay-inc3 & inc3-inc9 & inc9-inc17 & inc17-hatch & hatch-n5 & n5-n9\\
\hline
AR &  &  &  &  &  &  & \\
\hline
AVPR1B &  &  &  &  &  &  & \\
\hline
AVPR2 &  &  &  &  &  & FP(+) & \\
\hline
CRH &  &  &  &  &  &  & \\
\hline
CRHR1 &  &  &  &  &  &  & \\
\hline
DIO1 &  &  &  &  &  &  & \\
\hline
DIO2 &  & FG(-) &  &  &  &  & \\
\hline
DIO3 & FP(-) &  &  &  &  &  & \\
\hline
ESR2 &  &  &  &  &  &  & \\
\hline
FSHB &  &  &  &  &  &  & \\
\hline
FSHR &  &  &  &  &  &  & \\
\hline
GABRQ &  &  &  &  &  &  & \\
\hline
GALR1 &  &  &  &  &  &  & \\
\hline
GNRH1 &  &  &  &  &  &  & \\
\hline
GNRHR & FP(-) &  &  &  &  &  & \\
\hline
JAK2 &  &  &  &  &  &  & \\
\hline
LHCGR &  &  &  &  &  &  & \\
\hline
NPFFR1 &  &  &  &  &  &  & \\
\hline
NPVF &  &  &  &  &  &  & \\
\hline
NPY &  &  &  &  &  &  & \\
\hline
NR3C2 &  &  &  &  &  &  & \\
\hline
POMC &  &  &  &  &  &  & \\
\hline
PRL &  &  &  & FP(+); MP(+) &  &  & \\
\hline
PRLH & FP(-) & FP(+) &  & FP(+) &  &  & \\
\hline
PRLHR &  &  &  &  &  &  & \\
\hline
SERPINA4 &  &  &  &  &  &  & \\
\hline
VIPR1 &  &  &  &  &  &  & \\
\hline
\end{tabular}
\end{table}

@@

>>>>

We created the rock dove transcriptome by sequenced 987 RNA samples from
the hypothalamus, pituitary and gonads from the female and male of 166
rock dove pairs across (Fig. 1). These samples come from nine timepoints
that encapsulate important transitions in reproduction (Fig. 1A), four
offspring removal manipulations (Fig. 1B), and three offspring
replacement manipulations (Fig. 1C) to determine the influence of
internal mechanisms and external stimuli on gene expression.

Using T-distributed Stochastic Neighbor Embedding (tSNE) we show broad
patterns of variation due to treatment or sex in each tissue (Fig. 1D).
In the hypothalamus, tSNE dimension visually separates most of the
samples that were subject to an offspring manipulation from those from
naturally occurring parental stages. In the pituitary however, samples
cluster more by wether they were before incubation day 9 or after
incubation day 17. Neither the hypothalamus or pituitary show strong sex
differences at this level, but the gonads samples from very distinct,
sex-specific clusters.

@@

\caption{\label{tab:table1}Changes in candidate gene expression over the course of parental care in rock doves. Significant changes in gene expression are shown for X candidate genes. Positive and negative changes are denoted with (+) and (-) respectively for female (F) and male (M) hypothalamus (H), pituitary (P) and gonads (G). Comparisons are denoted in the column names. The first group is the reference group.}
\centering
\begin{tabular}[t]{l|l|l|l|l|l|l|l}
\hline
gene & bldg-lay & lay-inc3 & inc3-inc9 & inc9-inc17 & inc17-hatch & hatch-n5 & n5-n9\\
\hline
AVPR2 &  &  &  &  &  & FP(+) & \\
\hline
DIO2 &  & FG(-) &  &  &  &  & \\
\hline
DIO3 & FP(-) &  &  &  &  &  & \\
\hline
GNRHR & FP(-) &  &  &  &  &  & \\
\hline
PRL &  &  &  & FP(+); MP(+) &  &  & \\
\hline
PRLH & FP(-) & FP(+) &  & FP(+) &  &  & \\
\hline
\end{tabular}
\end{table}

@@

\caption{\label{tab:table2}Changes in candidate gene expression in response to removal. Significant changess in gene expression are shown for X candidate genes. Positive and negative changes are denoted with (+) and (-) respectively for female (F) and male (M) hypothalamus (H), pituitary (P) and gonads (G). Comparisons are denoted in the column names. The first group is the reference group.}
\centering
\begin{tabular}[t]{l|l|l|l|l}
\hline
gene & inc.d3\_m.inc.d3 & inc.d9\_m.inc.d9 & inc.d17\_m.inc.d17 & hatch\_m.n2\\
\hline
AVPR1B &  &  &  & FP(+)\\
\hline
AVPR2 &  &  &  & FP(+)\\
\hline
DIO3 &  &  & FP(+); MH(-) & FH(-); FP(+)\\
\hline
ESR2 &  &  & FH(-); MH(-) & \\
\hline
FSHB &  &  & FP(-) & FH(-); FP(-)\\
\hline
GABRQ &  &  & FH(+) & FH(+)\\
\hline
JAK2 &  &  & FP(+) & FP(+)\\
\hline
LHCGR &  &  & FH(-) & FH(-); MH(-)\\
\hline
NPVF & FH(-) &  & FH(-) & FH(-); MH(-)\\
\hline
NPY & FP(-) &  &  & MP(+)\\
\hline
PRL &  &  & FP(-) & MP(-)\\
\hline
PRLH &  &  &  & MH(+)\\
\hline
SERPINA4 &  &  & FP(+) & FP(+)\\
\hline
\end{tabular}
\end{table}

\caption{\label{tab:table3}Changes in candidate gene expression in response to replacement. Significant changess in gene expression are shown for X candidate genes. Positive and negative changes are denoted with (+) and (-) respectively for female (F) and male (M) hypothalamus (H), pituitary (P) and gonads (G). Comparisons are denoted in the column names. The first group is the reference group.}
\centering
\begin{tabular}[t]{l|l|l|l|l|l}
\hline
gene & inc.d9\_early & inc.d17\_prolong & hatch\_early & hatch\_prolong & hatch\_extend\\
\hline
AR &  &  & FP(+) &  & \\
\hline
AVPR2 &  &  & FP(+) &  & \\
\hline
CRHR1 &  & MG(+) &  &  & \\
\hline
ESR2 &  & MG(+) &  &  & \\
\hline
GABRQ & MP(+) & FH(+) & MH(+); MP(+) & FH(+); MG(+) & FH(+)\\
\hline
GNRH1 &  &  & FH(+) &  & \\
\hline
LHCGR &  & FH(-) & MH(-) & FH(-) & FH(-); MH(-)\\
\hline
NPVF & MH(-) & FH(-) & FH(-); MH(-) & FH(-) & FH(-)\\
\hline
NPY &  & MP(-) & MP(+) &  & \\
\hline
POMC &  &  & FH(-) & FH(-) & \\
\hline
PRL &  &  & FP(-); MP(-) &  & \\
\hline
PRLHR &  &  & MP(+) &  & \\
\hline
SERPINA4 &  & MG(+) & MH(+) &  & \\
\hline
\end{tabular}
\end{table}

levelsmanip

### 2020-09-08 manuscript/elife_v1.Rmd
now with 3 awesome tables

```{r setup, include = F}
library(tidyverse)
library(kableExtra)
library(cowplot)
```

@@

@@

```{r table1, echo =FALSE, message = F }
table1 <- read_csv("../results/table1.csv")
table1[is.na(table1)] <- " "
kable(table1,
      col.names = c('gene', 'bldg-lay',  'lay-inc3', 'inc3-inc9', 'inc9-inc17', 
                    'inc17-hatch', 'hatch-n5', 'n5-n9'),
      caption = "Changes in candidate gene expression. Significant chagns in gene expression are shown for X candidate genes. Positive and negative changes are denoted with (+) and (-) respectively for female (F) and male (M) hypothalamus (H), pituitary (P) and gonads (G). Comparisons are denoted in the column names. The first group is the reference group.") 
```

>>>>

```{r setup, include = F}
library(tidyverse)
library(kableExtra)
library(cowplot)

source("../R/themes.R")
```

@@

```{r cDEGs, include = F}
cDEGs <- read_csv("../results/03_wideDEGscandidates.csv") %>%
  select(-X1)

# helper function to add column of NAs if no DEGs exist
fncols <- function(data, cname) {
  add <-cname[!cname%in%names(data)]
  if(length(add)!=0) data[add] <- NA
  data
}

maketables123 <- function(df, whichlevels){
  table <- fncols(cDEGs, whichlevels) %>% 
    select(gene, whichlevels)  %>%
    mutate(allres = rowSums(is.na(.)))

  numcomparisons <- ncol(table) -2

  table <- table %>%
    filter(allres < numcomparisons) %>%
    select(-allres)
  table[is.na(table)] <- " "
  print(table)
  return(table)

}

table1 <- maketables123(cDEGs, levelssequential) 
table2 <- maketables123(cDEGs, levelsrm) 
table3 <- maketables123(cDEGs, levelsreplace)

write_csv(table1, "../results/table1.csv")
write_csv(table2, "../results/table2.csv")
write_csv(table3, "../results/table3.csv")

```

@@

```{r table1, echo =FALSE, message = F }
kable(table1,
      col.names = c('gene', 'bldg-lay',  'lay-inc3', 'inc3-inc9', 'inc9-inc17', 
                    'inc17-hatch', 'hatch-n5', 'n5-n9'),
      caption = "Changes in candidate gene expression over the course of parental care in rock doves. Significant changes in gene expression are shown for X candidate genes. Positive and negative changes are denoted with (+) and (-) respectively for female (F) and male (M) hypothalamus (H), pituitary (P) and gonads (G). Comparisons are denoted in the column names. The first group is the reference group.") 
```

```{r table2, echo =FALSE, message = F }
kable(table2,
      caption = "Changes in candidate gene expression in response to removal. Significant changess in gene expression are shown for X candidate genes. Positive and negative changes are denoted with (+) and (-) respectively for female (F) and male (M) hypothalamus (H), pituitary (P) and gonads (G). Comparisons are denoted in the column names. The first group is the reference group.") 
```

```{r table3, echo =FALSE, message = F }
kable(table3,
      caption = "Changes in candidate gene expression in response to replacement. Significant changess in gene expression are shown for X candidate genes. Positive and negative changes are denoted with (+) and (-) respectively for female (F) and male (M) hypothalamus (H), pituitary (P) and gonads (G). Comparisons are denoted in the column names. The first group is the reference group.") 
```

levelsmanip
